hbm_paper_results

### 2024-06-07 paper_main_results.Rmd
cleaning up

z.sex.var.plot <- ggplot(z.sex.var.plot.df) +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
             aes(fill = median_sex_diff)) +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-0.4, .4)) #
```

@@

z.sex.var.plot.global <- ggplot(z.sex.var.plot.global.df) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = median_sex_diff),
             side = c("coronal")) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-0.4, 0.4), name = "Male - Female Median \nCentile Error") #,
# sex.var.plot.global
```

@@

sex.var.plot.noext_z <- ggplot(sex.var.plot.df.noext_z) +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
             aes(fill = median_sex_diff)) +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits=c(-0.5, 0.5)) #
```

@@

noext.sex.var.plot.global_z <- ggplot(noext.sex.var.plot.global.df_z) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = median_sex_diff),
             side = c("coronal")) +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nCentile Error", limits = c(-0.5, 0.5)) #,
# sex.var.plot.global
```

@@

>>>>

z.sex.var.plot <- ggplot(z.sex.var.plot.df) +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
             aes(fill = median_sex_diff)) +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nZ-Score Error", limits = c(-0.4, .4)) #
```

@@

z.sex.var.plot.global <- ggplot(z.sex.var.plot.global.df) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = median_sex_diff),
             side = c("coronal")) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", limits = c(-0.4, 0.4), name = "Male - Female Median \nZ-score Error") #,
# sex.var.plot.global
```

@@

sex.var.plot.noext_z <- ggplot(sex.var.plot.df.noext_z) +
  theme_linedraw(base_size=12) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
             aes(fill = median_sex_diff)) +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nZ-score Error", limits=c(-0.5, 0.5)) #
```

@@

noext.sex.var.plot.global_z <- ggplot(noext.sex.var.plot.global.df_z) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = median_sex_diff),
             side = c("coronal")) +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Male - Female Median \nZ-score Error", limits = c(-0.5, 0.5)) #,
# sex.var.plot.global
```

@@

Redo with combat w/o covariates

```{r}
dataset_names <- c(raw = "Raw Data", cf = "ComBat w/o Covar", cf.lm = "ComBat", cf.gam ="ComBat-GAM", cf.gamlss ="ComBatLS")
#var plot
var.plot.df2 <- sigma.sex.df2 %>%
  dplyr::filter(pheno_cat != "Global Volume") %>%
  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE),
         sex_standard = ifelse(drop1.pval <0.05, sex_standard, NA)) %>%
  group_by(dataset, pheno_cat)

range(var.plot.df2$sex_standard, na.rm=T)

tp <- unique(var.plot.df2[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="raw")
tp$sex_standard <- 1

var.plot2 <- ggplot(var.plot.df2) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = dk_nomed, 
             position = position_brain(hemi ~ side),
             aes(fill = sex_standard)) +
  #scale_fill_gradient2() +
  facet_grid(dataset ~ pheno_cat, labeller = labeller(dataset = as_labeller(dataset_names))) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=2/3, plot.margin = margin(1,1,1,1, 'pt')) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
```

```{r}
#plot global
var.plot.global.df2 <- sigma.sex.df2 %>%
  dplyr::filter(pheno_cat == "Global Volume") %>%
  mutate(dataset = factor(dataset, levels = c("raw", "cf", "cf.lm", "cf.gam", "cf.gamlss"), ordered = TRUE)) %>%
  dplyr::select(!c(label, region, hemi)) %>% #drop cols corresponding to dk regions
  dplyr::mutate(tissue_class = case_when(pheno == "GMV" ~ "cGM",
                                            pheno == "Ventricles" ~ "CSF",
                                            pheno == "WMV" ~ "WM",
                                            pheno == "sGMV" ~ "sGM",
                                            TRUE ~ NA_character_),
         sex_standard = ifelse(drop1.pval <0.05, sex_standard, NA)) %>%
  group_by(dataset)

tp <- unique(var.plot.global.df2[,c('dataset','pheno_cat')]) %>%
  dplyr::filter(dataset=="raw")
tp$sex_standard <- 1

range(var.plot.global.df2$sex_standard, na.rm=T)

var.plot.global2 <- ggplot(var.plot.global.df2) +
  geom_rect(data = tp, fill="gray", xmin = -Inf,xmax = Inf,
            ymin = -Inf, ymax = Inf, alpha = 0.4) +
  ggseg::geom_brain(atlas = aseg_cerebrum, 
             aes(fill = sex_standard),
             side = c("coronal")) +
  #scale_fill_gradient2() +
  facet_grid(dataset~pheno_cat) + 
  theme_linedraw(base_size=12) +
  theme(axis.title = element_blank(), axis.text = element_blank(), axis.ticks = element_blank(), panel.grid.major = element_blank(), panel.grid.minor = element_blank(), panel.border = element_rect(color= "gray", fill = NA), aspect.ratio=, plot.margin = margin(1,1,1,1, 'pt'), strip.text.y = element_blank()) +
  paletteer::scale_fill_paletteer_c("grDevices::Cork", name = "Sex Effect", limits = c(-0.12, 0.12), breaks = c(-0.1, 0,0.1))
```

Combine plots
```{r}
p <- ggarrange(
  var.plot.global2, var.plot2,
  ncol=2, nrow=1, 
  widths = c(1, 3.18),
  #labels=c("A", "B"),
  #hjust=0.01,
  align = "hv",
  common.legend = TRUE, legend = "right"
  )
ggsave("figs/full_sex_eff_on_scale.jpeg", p, width=8, height=8, units="in")
```

##Test Sex-Effects Across Feature Types

```{r}
raw_sex_eff <- sigma.sex.df2 %>%
  dplyr::filter(dataset == "raw")

summary(aov(sex_standard ~ pheno_cat, data=raw_sex_eff))
pairwise.wilcox.test(raw_sex_eff$sex_standard, raw_sex_eff$pheno_cat, p.adjust.method = "fdr")
```

plot

```{r}
comp <- get_comparisons(raw_sex_eff, pheno_cat, ref.group = NULL)

? <- ggboxplot(raw_sex_eff, x = "pheno_cat", y = "sex_standard",
          color = "pheno_cat")+ 
  stat_compare_means(method = "wilcox.test", comparisons = comp, label="p.signif")+ # Add pairwise comparisons p-value
  stat_compare_means(label.y = 0.25) +     # Add global p-value
  theme(legend.position = "none") +
  labs(x="Brain Feature Category", y="Standardized Effect of Sex on Variance")

ggsave("figs/sex_eff_comp_plt.jpeg", sex_eff_comp_plt)
```

### 2024-06-07 paper_replication_results.Rmd
cleaning up

all_cent_sum_boxplt_wtest <- all_cent_sum.df %>%
  dplyr::filter(dataset != "raw") %>%
  mutate(source = factor(source, levels=c("low", "high"), ordered=TRUE)) %>%
  ggplot(aes(x=dataset, y=female_ratio)) +
  geom_boxplot(aes(color=dataset)) +
  # geom_dotplot(binaxis = "y", stackdir="center") +
  facet_grid(pheno_cat~source, scales="free_y", labeller = labeller(source=cent_ext_list)) +
  geom_hline(data=all_cent.sum_raw, aes(yintercept=female_ratio), linetype="dashed")+
  geom_text(aes(label = sig_fdr, y = max_val + 0.01),
              data = all_t_test.df) +
  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
  theme_linedraw() +
  scale_x_discrete(labels=dataset_names) +
  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y="Proportion Female", x = "Dataset")
#ggsave(filename="figs/all_cent_sum_boxplt_wtest.png", all_cent_sum_boxplt_wtest)
```

>>>>

all_cent_sum_boxplt.diff_full <- all_cent_sum.df_scaled %>%
  dplyr::filter(dataset != "raw") %>%
  mutate(source = factor(source, levels=c("low", "high", "ext"), ordered=TRUE)) %>%
  ggplot(aes(x=dataset, y=female_ratio_diff)) +
  geom_boxplot(aes(color=dataset)) +
  # geom_dotplot(binaxis = "y", stackdir="center") +
  facet_grid(pheno_cat~source, scales="free_y", labeller = labeller(source=cent_ext_list)) +
  geom_hline(yintercept=0, linetype="dashed")+
  geom_text(aes(label = sig_fdr, y = (y_pos)), nudge_y=.005,
               data = all_t_test.df) +
  scale_color_manual(values = c(cf= "#440D54", cf.gam = "#3D75C2", cf.lm = "#75C23D", cf.gamlss ="#CFB023FF")) + 
  theme_linedraw() +
  scale_x_discrete(labels=dataset_names) +
  # scale_y_continuous(labels=function(x)x/all_t_test.df$female_ratio[which(raw_df$pheno_cat == pheno_cat)]) + 
  theme(legend.position="none", plot.title = element_text(size=12), axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  labs(y="Female Bias", x = "Dataset")
ggsave(filename="figs/all_cent_sum_boxplt_wtest.png", all_cent_sum_boxplt.diff_full)
```
